---
title: "HW1"
author: "Haiyue Peng"
date: "2023-09-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1.1

First of all we load the data, and ask R to treat it as a categorical variable. 

```{r cars}
library(faraway)
data(teengamb)
teengamb$sex = factor(teengamb$sex)
```

Numerical summary of data: 

```{r}
summary(teengamb)
```

Graphical summary of data:

```{r}
par(mfrow=c(3,3), pin=c(1,0.8))
plot(teengamb$sex)
hist(teengamb$status, labels = TRUE)
boxplot(teengamb$status)
hist(teengamb$income, labels = TRUE)
boxplot(teengamb$income)
hist(teengamb$verbal, labels = TRUE)
boxplot(teengamb$verbal)
hist(teengamb$gamble, labels = TRUE)
boxplot(teengamb$gamble)
```

## Problem 1.2

From the numerical summary we know that

\begin{table}[ht]
\centering
\begin{tabular}{|c|c|c|}
\hline
 & \textbf{Mean} & \textbf{Median} \\
\hline
\textbf{income} & 4.642 & 3.250 \\
\textbf{gamble} & 19.3 & 6.0 \\
\hline
\end{tabular}
\end{table}

Their means are larger than medians. We can explain this through their histograms and box plots: these two variables are right skewed. There is a cluster of lower scores and a few exceptionally high scores. The few high scores tend to "pull" the mean towards them, making it larger than the median. 

## Problem 1.3

```{r}
uni <- unique(teengamb$verbal)
length(uni)
```
There are 9 different values in "verbal". 

Based on the box plot, we can consider those scores greater than 8 as outlying verbal scores. 

## Problem 1.4

```{r}
female <- teengamb[teengamb$sex==1, 3:5]
summary(female)
male <- teengamb[teengamb$sex==0, 3:5]
summary(male)

```
```{r}
par(mfrow=c(1,3))
hist(female$income, col=rgb(0,0,1,1/4), xlim=c(0,16), main="Income", freq = FALSE)
hist(male$income, col=rgb(1,0,0,1/4), xlim=c(0,16), add=T, freq = FALSE)
hist(female$verbal, col=rgb(0,0,1,1/4), xlim=c(0,10), main="Verbal", freq = FALSE)
hist(male$verbal, col=rgb(1,0,0,1/4), xlim=c(0,10), add=T, freq = FALSE)
hist(female$gamble, col=rgb(0,0,1,1/4), xlim=c(0,160), main="Gamble", freq = FALSE)
hist(male$gamble, col=rgb(1,0,0,1/4), xlim=c(0,160), add=T, freq = FALSE)
```

We can discover that female's income and gamble is lower than male's income and gamble, and that female's verbal data is more centralized than male's. 